---
title: "Logistic Regression Through Numerical Optimization"
author: "Final Group 4"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Package Overview

This R package contains functions designed to perform logistic regression on a user's chosen response and predictor data. This is done through estimating the coefficient vector $\beta$ and which contains response and predictor data and the intercept. This is done using the function
$$
\hat{\beta}\::=\underset{\beta}{\operatorname{argmin}}\sum_{i=1}^{n}(-y_{i}\cdot\ln(p_{i})-(1-y_{i})\cdot\ln(1-p_{i}))
$$
where
$$
p_{i}\::=\frac{1}{1+exp(-x_{i}^{T}\beta)}
$$
with $y_{i}$ representing the $i^{\text{th}}$ observation of the response and $x_{i}$ representing the $i^{\text{th}}$ row of the predictors.

#### ls_optim
The least squares optimizer uses a standard approach to a regression analysis. The ls_optim function works on finding how gender affects income.
```{r ls_optim, include = FALSE, echo = TRUE}

ls_optim(gender, income)

```

#### bootstrap_ci
The bootstrap intervals helps to focus on random sampling with replacement.
```{r bootstrap_ci, include = FALSE, echo = TRUE}

bootstrap_ci(gender, income, alpha = 0.05, rounds = 20)

```


#### conf_matrix
The confusion matrix shows the performance of the ls_optim algorithm in a table. 
```{r conf_matrix, include = FALSE, echo = TRUE}

conf_matrix(gender, income, alpha = 0.05, cutoff = 0.5)

```

#### log_curve
The log curve uses the ls_optim function to demonstrate a logistic curve in a plot.
```{r log_curve, include = FALSE, echo = TRUE}
log_curve(income, gender)
```

#### plot_metrics
The plot metrics uses the conf_matrix function to show the plot of accuracy in the confusion metrics. 
```{r plot_metrics, include = FALSE, echo = TRUE}

plot_metrics(gender, income, alpha = 0.05)

```

## Example: ___ Data
